---
title: "Preprosim"
author: "Markus Vattulainen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Preprosim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Data quality simulation can be used to check the robustness of data
analysis findings and learn about the impact of data quality contaminations on
classification. This package helps to add contaminations (noise, missing values,
outliers, low variance, irrelevant features, class swap (inconsistency), class
imbalance and decrease in data volume) to data and then evaluate the simulated
data sets for classification accuracy. As a lightweight solution simulation runs
can be set up with no or minimal up-front effort.

## Quick start

Example 1 (not run):

```{r}
# library(preprosim)
# preprosimplot(preprosimrun(iris))
```
Boxplot of classification accuracies. 6561 datasets with all eight contaminations
each having three contamination intensities. Classification accuracies computed 
with stochastic gradient boosting (gbm) and holdout validated 10 times. 
Runtime several hours. 


Example 2:

Checking the impact of missing values (primary) and noise (secondary) on
classification accuracy. Iris data, rpart model and two times holdout
validation.
 

```{r, results='hide', message=FALSE, warning=FALSE, fig.width=8}
library(preprosim)
res <- preprosimrun(iris, param=newparam(iris, "custom", x="misval", z="noise"), caretmodel="rpart", holdoutrounds = 2, verbose=FALSE)
preprosimplot(res)
preprosimplot(res, "xz", x="misval", z="noise")
```

Adding noise and missing values has a substantial effect on classification accuracy (top).
The relationship between missing values and classification accuracy in similar in the three
categories of noise (bottom).
